---
title: "Untitled"
author: "Mark Dunning"
date: "10 February 2016"
output: html_document
---

[Transcriptional profiling of the epigenetic regulator Smchd1](http://www.sciencedirect.com/science/article/pii/S2213596015301306)

Details are given in the [Supplementary Materials](http://bioinf.wehi.edu.au/folders/smchd1/Smchd1.html)

```{r}
library(SRAdb)
sqlfile <-'SRAmetadb.sqlite'
if(!file.exists('SRAmetadb.sqlite')) sqlfile <<- getSRAdbFile()
```

```{r}
sra_con <- dbConnect(SQLite(),sqlfile)
```

```{r}
sraInf <- getSRAinfo("SRP051084",sra_con, sraType="sra")
sraInf
```


```{r cache=TRUE}
sapply(sraInf$run, function(x) getSRAfile(x,sra_con, fileType="fastq"))
```

Next, download the *mm10* reference sequence from UCSC.

```{eval=FALSE}
wget --timestamping 'ftp://hgdownload.cse.ucsc.edu/goldenPath/mm10/bigZips/chromFa.tar.gz' -O chromFa.tar.gz
gunzip chromFa.tar.gz
tar xvf chromFa.tar
cat *.fa > mm10.fa
rm chr*.fa
```

It is a good idea to use the [fastqc](http://www.bioinformatics.babraham.ac.uk/projects/fastqc/) tool to perform a QA on these files. We can do this at the command-line using the following loop.

```{}
for file in *.fastq.gz
do
fastqc $file
done
```

```{r}
fastqc_reports <- list.files(pattern="fastqc.zip")
sapply(fastqc_reports,unzip)
```

Now we can use the `Rsubread` package to build an index for the single reference fasta file that we have just created.

```{r cache=TRUE}
library(Rsubread)
buildindex("mm10",reference = "mm10.fa")
```

Wild-Type replicate 1 (GSM1564327) appears to be run across four single-end runs

```{r}
align(readfile1="SRR1706957.fastq.gz",index = "mm10")
align(readfile1="SRR1706958.fastq.gz",index = "mm10")
align(readfile1="SRR1706959.fastq.gz",index = "mm10")
align(readfile1="SRR1706960.fastq.gz",index = "mm10")
```

We can merge the bam files into one using `Rsamtools`.

```{r cache=TRUE}
library(Rsamtools)
fls <- c("SRR1706957.fastq.gz.subread.BAM","SRR1706958.fastq.gz.subread.BAM","SRR1706959.fastq.gz.subread.BAM","SRR1706960.fastq.gz.subread.BAM")
mergeBam(fls, destination ="GSM1564327.bam")
sortBam(file ="GSM1564327.bam",destination ="GSM1564327.sorted")
indexBam("GSM1564327.sorted.bam")
```


GSM1564328: Smchd1 null replicate 1

N.B. alignment took around 50 minutes per-sample

```{r cache=TRUE}
align(readfile1 = "SRR1706961_1.fastq.gz", readfile2 = "SRR1706961_2.fastq.gz",index="mm10",output_file = "GSM1564328.bam")
```

 GSM1564329: Smchd1 null replicate 2

```{r cache=TRUE}
align(readfile1 = "SRR1706962_1.fastq.gz", readfile2 = "SRR1706962_2.fastq.gz",index="mm10",output_file = "GSM1564329.bam")
```

GSM1564330: Smchd1 null replicate 3

```{r cache=TRUE}
align(readfile1 = "SRR1706963_1.fastq.gz", readfile2 = "SRR1706963_2.fastq.gz",index="mm10",output_file = "GSM1564330.bam")
```

GSM1564331: Wild-type replicate 2

```{r cache=TRUE}
align(readfile1 = "SRR1706964_1.fastq.gz", readfile2 = "SRR1706964_2.fastq.gz",index="mm10",output_file = "GSM1564331.bam")
```

GSM1564332: Wild-type replicate 3

```{r cache=TRUE}
align(readfile1 = "SRR1706965_1.fastq.gz", readfile2 = "SRR1706965_2.fastq.gz",index="mm10",output_file = "GSM1564332.bam")
```

GSM1564333: Smchd1 null replicate 4

```{r cache=TRUE}
align(readfile1 = "SRR1706966_1.fastq.gz", readfile2 = "SRR1706966_2.fastq.gz",index="mm10",output_file = "GSM1564333.bam")
```

## Counting


```{r}
library(GEOquery)
geoInf <- getGEO("GSE64099")
targets <- pData(geoInf[[1]])
```


```{r}
library(Rsubread)
fls <- paste0(rownames(targets), ".sorted.bam")
all(file.exists(fls))

counts <- featureCounts(fls, annot.inbuilt = "mm10")
save(counts, file="lymphoma-counts.rda")
```



## Differential Expression

(from online markdown doc)

```{r}


load("lymphoma-counts.rda")
library(org.Mm.eg.db)
geneanno <- select(org.Mm.eg.db, keys = as.character(counts$annotation$GeneID), keytype = "ENTREZID",columns=c("SYMBOL"))
geneanno <- data.frame(geneanno, counts$annotation)

library(edgeR)
library(limma)

x <- list()
x$counts <- counts$counts
x$genes <- geneanno

sel = rowSums(cpm(x$counts)>0.5)>=3

x$counts = x$counts[sel,]
x$genes = x$genes[sel,]
x$anno = x$anno[sel,]

samplenames = colnames(x$counts)
genotype = ifelse(targets$characteristics_ch1.3 == "genotype: Smchd1 Wild-type, Eµ-Myc transgenic male","WT","Smcdh1-null")
sampleanno = data.frame("sampleID"=samplenames, "genotype"=genotype, "group"=genotype)

x$samples = sampleanno
x$samples$lib.size = colSums(x$counts)
x$samples$norm.factors = 1

row.names(x$samples) = colnames(x$counts)

x = new("DGEList", x)
dim(x)


```

```{r}
des = model.matrix(~genotype)
des[,2]=1-des[,2]
colnames(des)[2] = "WT"


x = calcNormFactors(x, method="TMM")
x = estimateDisp(x, robust=TRUE)
```

```{r}
sqrt(x$common.disp)
```

```{r}
v=voomWithQualityWeights(x, design=des, normalization="none", plot=TRUE,col=c("deepskyblue" ,"deepskyblue" ,"deepskyblue","deepskyblue", "darkolivegreen4", "darkolivegreen4","darkolivegreen4"))
```

```{r}
vfit = lmFit(v,des)
vtfit=treat(vfit,lfc=log2(1.5))
vfit= eBayes(vfit)
results <- decideTests(vfit,p.value=0.01)
summary(results)
topTable(vfit)
```

